---
title: "TSF"
output: html_notebook
---

# Overview
- concepts in ts
- descriptive analysis ts
- exponential smoothing
- ARIMA
- rstudio

# UV TSA
- Exponential Smoothing
- ARIMA
- ARCH Models

# MV TSA
- Vector Autoregressive Models
- Error Correction Models
- Cointegration

Setting up working directory

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/R/PSAI_Training_2018/TSF/MODULE 3- TRAINING ON TIME SERIES ANALYSIS AND FORECASTING/day 1/Datasets/Datasets')
```

note that there are problems in running head function

```{r}
inflation <- read.table("inflation.txt",header = T)
inflation$month <- gsub("M","-",inflation$month)
month <- zoo::as.yearmon(inflation$month, format="%Y-%m") #zoo package
inflation <- cbind(month,inflation["inflation"])
plot(inflation)
plot(inflation, type="o")
plot(inflation, type="l",col="blue",xlab="",ylab="Inflation Rate",
                main="Inflation Rate \n(Jan 2013 - Sep 2018)")
```

Hodrick-Prescott Filter

```{r}
inflation.hp <- mFilter::hpfilter(inflation$inflation, freq=14400, type="lambda")
plot(inflation.hp)
inflation <- cbind(inflation,inflation.hp$trend)
colnames(inflation)[3] <- "hp_trend"
```

ggplot input: Dataframe

```{r}
library(ggplot2)
ggplot(data=inflation, aes(x=month)) + 
  geom_line(aes(y=inflation, colour="Inflation Rate")) + 
  geom_line(aes(y=hp_trend, colour="HP Trend")) +
  labs(x="Month", y="Inflation Rate", colour="") +
  theme(legend.position="top") +
  scale_colour_manual(values=c("blue","red"))
```

Workshop 1

```{r}
rgdp.ts<-ts(read.table("rgdp2.txt",header = T),frequency=4,start=c(1998,1),end=c(2018,2))

#line plot with elements
plot(rgdp.ts, main="QuartelyGDP of the Philippines: Q1 1998 -Q2 2018
(at constant 2000 prices)",
ylab="Real GDP (in Million Pesos)", xlab="")
```

Workshop 1

```{r}
#time series decomposition
decomp_rgdp<-decompose(rgdp.ts,"additive")
#plot components
par(mfrow=c(2,2)) #run this par if you want a single panel of four graphs
plot(rgdp.ts,ylab="Observed",xlab="")
plot(decomp_rgdp$trend,ylab="Trend-Cycle",xlab="")
plot(decomp_rgdp$seasonal,ylab="Seasonality",xlab="")
plot(decomp_rgdp$random,ylab="Irregular",xlab="")
```

Workshop 2

```{r}
export <- read.table("export.txt",header = T)
export$month <- gsub("M","-",export$month)
month <- zoo::as.yearmon(export$month, format="%Y-%m") #zoo package
export <- cbind(month,export["export"])
plot(export, type="l")
```


```{r}
export.hp <- mFilter::hpfilter(export$export, freq=14400, type="lambda")
plot(export.hp)
export <- cbind(export,export.hp$trend)
colnames(export)[3] <- "hp_trend"
```



```{r}
library(ggplot2)
ggplot(data=export, aes(x=month)) + 
  geom_line(aes(y=export, colour="Exports")) + 
  geom_line(aes(y=hp_trend, colour="HP Trend")) +
  labs(x="Month", y="Exports", colour="") +
  theme(legend.position="top") +
  scale_colour_manual(values=c("blue","red"))
```

# Exponential Smoothing (no minimum requirement)

ANN - no trend no seas  
AAdN -  dampened trend no seas  
AAN - w/trend no seas  
AAA - w/trend add seas  
AAM - w/trend, mult seas

```{r}
export <- read.table("export.txt", header=T)
export$month <- gsub("M","-",export$month)
rtime <- zoo::as.yearmon(export$month, format="%Y-%m")
export <- cbind(rtime,export["export"])
export.ts <- ts(export$export,frequency=12,start=c(1991,1))
plot.ts(export.ts)
```

```{r}
library(smooth)
export.ses  <- es(export.ts,model="ANN",h=12,holdout=FALSE,intervals="parametric",level=0.95,cfType="MSE")
export.des  <- es(export.ts,model="AAdN",h=12,holdout=FALSE,intervals="parametric",level=0.95,cfType="MSE")
export.hwn  <- es(export.ts,model="AAN",h=12,holdout=FALSE,intervals="parametric",level=0.95,cfType="MSE")
export.hwa  <- es(export.ts,model="AAA",h=12,holdout=FALSE,intervals="parametric",level=0.95,cfType="MSE")
export.hwm  <- es(export.ts,model="AAM",h=12,holdout=FALSE,intervals="parametric",level=0.95,cfType="MSE")
export.auto <- es(export.ts,model="ZZZ",h=12,holdout=FALSE,intervals="parametric",level=0.95,cfType="MSE")
```

ANN  means Additive errors, No trend, No seasonality (SES)  
AAdN means Additive errors, Additive (dampled) trend, No seasonality (DES)  
AAN  means Additive errors, Additive trend, No seasonality (Holt-Winters no seasonality)  
AAA  means Additive errors, Additive trend, Additive seasonality (HW with additive seasonality)  
AAM  means Additive errors, Additive trend, Multiplicative seasonality (HW with multiplicative seasonality)  
ZZZ  means all components are (Z) estimated  

```{r}
MAPE(export.ts,export.ses$fitted)
MAPE(export.ts,export.des$fitted)
MAPE(export.ts,export.hwn$fitted)
MAPE(export.ts,export.hwa$fitted)
MAPE(export.ts,export.hwm$fitted)
MAPE(export.ts,export.auto$fitted)
```

```{r}
summary(export.auto)
```


```{r}
#Forecast h steps using model with lowest MAPE
forecast(export.auto,h=24)
plot(forecast(export.auto,h=24))
```

Workshop

```{r}
rgdp.ses  <- es(rgdp.ts,model="ANN",h=8,holdout=FALSE,intervals="parametric",level=0.95,cfType="MSE")
rgdp.des  <- es(rgdp.ts,model="AAdN",h=8,holdout=FALSE,intervals="parametric",level=0.95,cfType="MSE")
rgdp.hwn  <- es(rgdp.ts,model="AAN",h=8,holdout=FALSE,intervals="parametric",level=0.95,cfType="MSE")
rgdp.hwa  <- es(rgdp.ts,model="AAA",h=8,holdout=FALSE,intervals="parametric",level=0.95,cfType="MSE")
rgdp.hwm  <- es(rgdp.ts,model="AAM",h=8,holdout=FALSE,intervals="parametric",level=0.95,cfType="MSE")
rgdp.auto <- es(rgdp.ts,model="ZZZ",h=8,holdout=FALSE,intervals="parametric",level=0.95,cfType="MSE")
```

```{r}
MAPE(rgdp.ts,rgdp.ses$fitted)
MAPE(rgdp.ts,rgdp.des$fitted)
MAPE(rgdp.ts,rgdp.hwn$fitted)
MAPE(rgdp.ts,rgdp.hwa$fitted)
MAPE(rgdp.ts,rgdp.hwm$fitted)
MAPE(rgdp.ts,rgdp.auto$fitted)
```

```{r}
#Forecast h steps using model with lowest MAPE
forecast(rgdp.auto,h=8)
plot(forecast(rgdp.auto,h=8))
```

```{r}
sales <- read.table("sales.txt", header=T)
sales$month <- gsub("Q","-",sales$quarter)
rtime <- zoo::as.yearqtr(sales$quarter, format="%Y-%q")
sales <- cbind(rtime,sales["sales"])
sales.ts <- ts(sales$sales,frequency=4,start=c(2001,2))
plot.ts(sales.ts)
```

```{r}
sales.ses  <- es(sales.ts,model="ANN",h=8,holdout=FALSE,intervals="parametric",level=0.95,cfType="MSE")
sales.des  <- es(sales.ts,model="AAdN",h=8,holdout=FALSE,intervals="parametric",level=0.95,cfType="MSE")
sales.hwn  <- es(sales.ts,model="AAN",h=8,holdout=FALSE,intervals="parametric",level=0.95,cfType="MSE")
sales.hwa  <- es(sales.ts,model="AAA",h=8,holdout=FALSE,intervals="parametric",level=0.95,cfType="MSE")
sales.hwm  <- es(sales.ts,model="AAM",h=8,holdout=FALSE,intervals="parametric",level=0.95,cfType="MSE")
sales.auto <- es(sales.ts,model="ZZZ",h=8,holdout=FALSE,intervals="parametric",level=0.95,cfType="MSE")
```

```{r}
MAPE(sales.ts,sales.ses$fitted)
MAPE(sales.ts,sales.des$fitted)
MAPE(sales.ts,sales.hwn$fitted)
MAPE(sales.ts,sales.hwa$fitted)
MAPE(sales.ts,sales.hwm$fitted)
MAPE(sales.ts,sales.auto$fitted)
```

```{r}
#Forecast h steps using model with lowest MAPE
forecast(sales.des,h=8)
plot(forecast(sales.des,h=8))
```

Day 2

ARIMA

Investigate the "dependency structure".

- how the past values affect the present and future values.

Unit Root Test (Stationarity Tests)

- The presence of unit root makes it non-stationarity.

How to detect unit roots.

1. line graphs (suggestive)
2. Correlogram (suggestive)
3. unit root tests (conclusive)

ADF Test Setup

$H_0:$ The series has a unit root. (non stationary)  
$H_a:$ The series has no unit root. (stationary)

Load Packages

```{r}
library(forecast)
library(tseries)
```

```{r}
#Performing an ADF test on Peso-Dollar Excahange data
pder <- read.table("pder.txt", header=T)
head(pder)
```

```{r}
pder.ts <- ts(pder$pder,frequency=12,start=c(1980,1))
head(pder.ts)
```

```{r}
plot(pder.ts)
```

```{r}
plot(diff(pder.ts))
```

```{r}
adf.test(pder.ts, alternative="stationary")
adf.test(diff(pder.ts), alternative="stationary")
```

Autocorrelation

- strength of linear correlation with it self.

Partial Autocorrelation

- removing effects.

ACF

\[
Corr(y_t,y_{t+k})
\]

PACF

\[
Corr(y_t,y_{t-1}|y_{t-1},y_{t-2},\cdots,y_{t-k+1})
\]

```{r}
#CPI Data
cpi <- read.table("cpi.txt", header=T)
cpi.ts <- ts(cpi$cpi,frequency=12,start=c(1994,1))
```

```{r}
adf.test(cpi.ts, alternative="stationary")
adf.test(diff(cpi.ts), alternative="stationary")
```

```{r}
par(mar = c(4, 4, 4, 1))
plot(cpi.ts)
acf(cpi.ts)
pacf(cpi.ts)
```

```{r}
par(mar = c(4, 4, 4, 1))
plot(diff(cpi.ts))
acf(diff(cpi.ts))
pacf(diff(cpi.ts))
```

```{r}
#CPI Data
rain <- read.table("rain.txt", header=T)
rain.ts <- ts(rain$ave_rain,frequency=12,start=c(2009,1))
```

```{r}
adf.test(rain.ts, alternative="stationary")
adf.test(diff(rain.ts), alternative="stationary")
```

```{r}
par(mar = c(4, 4, 4, 1))
plot(rain.ts)
acf(rain.ts)
pacf(rain.ts)
```

```{r}
par(mar = c(4, 4, 4, 1))
plot(diff(rain.ts))
acf(diff(rain.ts))
pacf(diff(rain.ts))
```

Autoregressive Process

Process - Behavior

Characterizing the ts (ACF, PACF)

$AR(p)$ ACF tapers off, PACF cuts off after lag $p$

$MA(q)$ ACF cuts off after lag $q$. PACF tapers off

$ARMA(p,q)$ both ACF and PACF tapers off  
challege: know the *correct* combination of $p$ and $q$

Workshop

```{r}
par(mar = c(4, 4, 4, 1))
acf(diff(pder.ts))
pacf(diff(pder.ts))
```

$MA(1)$ since the ACF cuts off at lag 1.

$AR(1)$ since the PACF cuts off at lag 1.

```{r}
par(mar = c(4, 4, 4, 1))
plot(rgdp.ts)
acf(rgdp.ts)
pacf(rgdp.ts)
```

```{r}
adf.test(rgdp.ts, alternative="stationary")
```

```{r}
par(mar = c(4, 4, 4, 1))
plot(diff(rgdp.ts,1))
acf(diff(pder.ts))
pacf(diff(pder.ts))
```

```{r}
adf.test(diff(rgdp.ts), alternative="stationary")
```




